Waiting time analysis

### 2025-04-25 notebooks/02_waiting_time.qmd
docs: update

using CairoMakie
```

```{julia}
wind_df = load(; name="Wind", ts=missing);
jno_df = load();
```

@@

get_fit_params(x; dist=Weibull, func=mean) = func(fit(dist, x))
get_wt_fit_params(time; kwargs...) = get_fit_params(waiting_time(time); kwargs...)
```

```{julia}
jno_ρ_df = @chain jno_df begin
    @transform!(
        :time_bin = (:time - minimum(:time)) ./ time_interval .|> round,
        :l = :L_k .* (abs ∘ cosd).(:θ_vn),
        :l_n = :L_k .* (abs ∘ cos ∘ atan).(:Vn_y ./ :Vn_x)
    )
end;
```

>>>>

using Discontinuity: remove_duplicates
using CairoMakie
```

```{julia}
df = load_taus(; unique_f=missing)

wind_df = load(; name="Wind", ts=missing);
jno_df = load();
```

@@

```{julia}
using Distributions
import DataFrames: groupby

get_fit_params(x; dist=Weibull, func=mean) = func(fit(dist, x))
get_wt_fit_params(time; kwargs...) = get_fit_params(waiting_time(time); kwargs...)

function groupby(df::AbstractDataFrame, dt::Period; col=:time)
    @chain df begin
        @transform!(:time_bin = round.(:time, dt),)
        groupby(:time_bin)
    end
end

begin
    dt = Day(50)
    tdf = @chain df begin
        remove_duplicates
        groupby(dt)
        combine(
            nrow,
            :time => get_wt_fit_params => :τ,
            [:r] .=> mean
        )
        @transform(:ρ = 1 ./ :τ)
    end

    x = :time_bin => "Time"
    specs = data(tdf) * mapping(x, :ρ => "Occurence rates")
    draw(specs)
end
```

### Normalization by thickness

### 2025-04-25 notebooks/21_model.qmd
docs: update

```{julia}
using DrWatson
using Discontinuity
using Discontinuity: Alfven_current
using DataFramesMeta
using PlasmaFormulary: inertial_length

@@

```{julia}
using CSV
using DataFrames

function HAPIDateTime(t::AbstractString)
    # Check if the string ends with Z and remove it for parsing
    t = rstrip(t, ['Z'])
    # Split into date and time parts if T is present
    parts = split(t, "T", limit=2)
    date_part = parts[1]
    time_part = length(parts) > 1 ? parts[2] : ""

    # Parse date part
    date = @something(
        tryparse(Date, date_part),
        tryparse_doy_date(date_part),
    )

    # Create DateTime with or without time
    if isempty(time_part)
        dt = DateTime(date)
    else
        dt = date + Time(time_part)
    end

    return dt
end

"""
Parse a date in Day of Year format (YYYY-DDD)
"""
function tryparse_doy_date(date_str::AbstractString)
    doy_regex = r"^(\d{4})-(\d{3})$"

    if occursin(doy_regex, date_str)
        m = match(doy_regex, date_str)
        year = parse(Int, m[1])
        doy = parse(Int, m[2])
        return Date(year) + Day(doy - 1)
    end

    return nothing
end

@@

Reads the solar wind parameters CSV file and returns a DataFrame.
"""
function read_sw_params_csv(file=datadir("jgra54158-sup-0001-supinfo.csv"))
    df = CSV.read(file, DataFrame)
    df[!, :UTC] = HAPIDateTime.(String.(df[!, :UTC]))
    return df
end

jno_sw_params = read_sw_params_csv()

@@

@chain jno_df_with_params begin
    @transform! begin
        :N_PROTONS_CC = :N_PROTONS_CC .* u"cm^-3"
        :d_star = :d_star .* u"nT/s"
        :duration = :duration .* u"s"
        :V_KMPS = :V_KMPS .* u"km/s"
    end
    @transform! begin
        :V_n_cross_sw = @. abs(first(:n_cross) * :V_KMPS)
        :d_i_sw = inertial_length.(:N_PROTONS_CC, Unitful.q, Unitful.mp)
        :j_Alfven_sw = Alfven_current.(:b_mag, :N_PROTONS_CC)
    end
    @transform! begin
        :L_n_cross_sw = :V_n_cross_sw .* :duration
        :j0_k_sw = @. abs(:d_star / :V_n_cross_sw / μ0) |> u"nA/m^2"
    end
    @transform! begin
        :L_n_cross_sw_norm = @. :L_n_cross_sw / :d_i_sw |> NoUnits
        :j0_k_sw_norm = @. :j0_k_sw / :j_Alfven_sw |> NoUnits
    end
end
```

>>>>

```{julia}
using DrWatson
using Discontinuity
using Discontinuity: Alfven_current
using DataFramesMeta
using PlasmaFormulary: inertial_length
using SpaceDataModel: tryparse_datetime

@@

```{julia}
using CSV
using DataFrames
using Unitful

@@

Reads the solar wind parameters CSV file and returns a DataFrame.
"""
function read_sw_params_csv(file=datadir("jgra54158-sup-0001-supinfo.csv"))
    df = CSV.read(file, DataFrame)
    @chain df begin
        @transform! @byrow begin
            :time = tryparse_datetime(String(:UTC))
            :N_PROTONS_CC = :N_PROTONS_CC * u"cm^-3"
            :V_KMPS = :V_KMPS * u"km/s"
        end
    end
    return df
end

@b read_sw_params_csv()

@@

@chain jno_df_with_params begin
    @transform! begin

        :d_star = :d_star .* u"nT/s"
        :duration = :duration .* u"s"
        :V_KMPS = :V_KMPS .* u"km/s"
    end
    @transform! begin
        :V_n_cross_sw = @. abs(first(:n_cross) * :V_KMPS)
        :d_i_sw = inertial_length.(:N_PROTONS_CC, Unitful.q, Unitful.mp)
        :j_Alfven_sw = Alfven_current.(:b_mag, :N_PROTONS_CC)
    end
    @transform! begin
        :L_n_cross_sw = :V_n_cross_sw .* :duration
        :j0_k_sw = @. abs(:d_star / :V_n_cross_sw / μ0) |> u"nA/m^2"
    end
    @transform! begin
        :L_n_cross_sw_norm = @. :L_n_cross_sw / :d_i_sw |> NoUnits
        :j0_k_sw_norm = @. :j0_k_sw / :j_Alfven_sw |> NoUnits
    end
end
```
